---
title: "LCart"
author: "Akshay Amrit"
date: "14/11/2019"
output: html_document
---

## Data

Loading Data
```{r}
library(readxl)
LCart <- read_excel("~/Business Case_Lendingkart (1).xlsx",
                    sheet = "Data", col_types = c("date",
                                                  "numeric", "numeric", "text", "text",
                                                  "text", "text", "text", "numeric",
                                                  "numeric", "text", "numeric", "numeric"))
```


## Task

- What were the total sales for each category in Wk 29?
- Share the sales and margin trends by category over the last 4 weeks basis the total orders, cost and margins. Create charts to depict the trends
- What were the Top 5 selling products in terms of  units over the 4 weeks and what were their margin (in % and $)?
- Which category had the best sales growth over the 4 weeks?
- Which category had the worst sales growth over the 4 weeks, and what would you do to improve performance?
- What was the average order value in Electronics over the 4 weeks? How could this be increased?
- Identify the Top 10 selling item types by value and units, and propose 3 ideas to grow sales even more on them
- List the Top 5 and Bottom 5 sellers by value over the 4 weeks, and show their sales by week.
- Highlight the best and worst thing about the business and what are the areas we can improve in.
- Give 3 months sales projection for top 3 categories with the suggestive actions to achieve the same
- Provide any other insights that you can find from the data source


## Data Cleaning

```{r}
library(tidyverse)
colnames(LCart) <- str_replace_all(colnames(LCart), " ", ".")
LCart$Order.Date <- strptime(LCart$Order.Date, format = "%Y-%m-%d")
LCart$Order.Date <- format(LCart$Order.Date, "%d/%m/%Y")
LCart$`Margin.%` <- str_replace_all(LCart$`Margin.%`, "%", "")
LCart$`Margin.%` <- as.numeric(LCart$`Margin.%`)
```
## Data Analysis

```{r}
str(LCart)

# No of duplicate Order ID
nrow(LCart) - length(unique(LCart$Order.ID))

```

## Task 1
What were the total sales for each category in Wk 29?
```{r}
LCart %>% filter(Week.Number == 29) %>% 
      group_by(Category) %>% 
      summarize(Sales.Per.Cat = sum(Price))
```

## Task 2
Share the sales and margin trends by category over the last 4 weeks basis the total orders, cost and margins. Create charts to depict the trends
```{r}
TwoTable <- LCart %>% 
      filter(Week.Number > 27) %>% 
      group_by(Category, Week.Number) %>% 
      summarize(Sales = sum(Quantity.sold), Margin = sum(`Margin.%`/100 * Cost))
g <- ggplot(data = TwoTable)
g + geom_line(aes(x = Week.Number, y = Sales, group = Category, colour = Category)) + 
      geom_point(aes(x = Week.Number, y = Sales, group = Category, colour = Category))
g + geom_line(aes(x = Week.Number, y = Margin, group = Category, colour = Category)) + 
      geom_point(aes(x = Week.Number, y = Margin, group = Category, colour = Category))
```


## Task 3
What were the Top 5 selling products in terms of  units over the 4 weeks and what were their margin (in % and $)?
```{r}
LCart <- LCart %>% mutate(Margin = Price - Cost)
Table.Three <- LCart %>%
      filter(Week.Number > 27) %>%
      group_by(Item.Title,Bar.Code) %>%
      summarize(Quantity = sum(Quantity.sold),TotalMargin = sum(Margin), TotalCost = sum(Cost), Margin.Perc = (TotalMargin / TotalCost) * 100) %>%
      arrange(desc(Quantity))
head(as.data.frame(Table.Three), n = 5)
```

## Task 4
Which category had the best sales growth over the 4 weeks?
```{r}
Table.Four <- LCart %>%
      filter(Week.Number > 27) %>%
      group_by(Category, Week.Number) %>%
      summarize(Sales = sum(Price)) %>%
      spread(Week.Number, Sales)
Table.Four$`28` <- replace_na(Table.Four$`28`, 0)
colnames(Table.Four) <- c("Category", "Week1", "Week2", "Week3", "Week4")
Table.Four <- Table.Four %>% group_by(Category) %>% mutate(Sales.Growth = Week4 - Week1, Growth.Margin = Sales.Growth/Week1 * 100) %>% arrange(desc(Sales.Growth))
Table.Four[1,]
```

## Task 5
Which category had the worst sales growth over the 4 weeks, and what would you do to improve performance?
```{r}
Table.Four[nrow(Table.Four),]
```

## Task 6
What was the average order value in Electronics over the 4 weeks? How could this be increased?
```{r}
LCart %>% filter(Week.Number > 27) %>% 
      group_by(Order.ID, Category) %>% 
      summarize(Order.Value = sum(Price)) %>%
      group_by(Category) %>%
      summarize(Average.Order.Value = mean(Order.Value)) %>%
      filter(Category == "Electronics")
```

## Task 7
Identify the Top 10 selling item types by value and units, and propose 3 ideas to grow sales even more on them
```{r}
Table.Seven <- LCart %>% group_by(Item.Type) %>%
      summarize(Value = sum(Price), Units = sum(Quantity.sold))

#Top 10 selling Item Type by Value
head(Table.Seven %>% arrange(desc(Value)), n = 10)

#Top 10 selling Item Type by Units
head(Table.Seven %>% arrange(desc(Units)), n = 10)
```

## Task 8
List the Top 5 and Bottom 5 sellers by value over the 4 weeks, and show their sales by week
```{r}
Table.Eight <- LCart %>%
      filter(Week.Number > 27) %>%
      group_by(Seller.Name) %>%
      summarize(Value = sum(Price))

# Top 5 sellers by Value
head(Table.Eight %>% arrange(desc(Value)), n = 5)

# Top 5 sellers by Value
head(Table.Eight %>% arrange(Value), n = 5)
```

## Task 9
Highlight the best and worst thing about the business and what are the areas we can improve in.  
?????