---
title: "Central Limit Theorem"
author: "Akshay Amrit"
date: "25th June, 2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objective  
This is a web page presentation created using R Markdown. This presentation is created 
to demonstrate my understanding of the topic. Being a very important concept which 
is also often not well understood, I will be demonstrating the concept visually 
using the shiny package.

  
#### Definition:
Central Limit Theorem. The central limit theorem states that if you have a population
with mean μ and standard deviation σ and take sufficiently large random samples from 
the population with replacement , then the distribution of the sample means will be 
approximately normally distributed.


## Process
User can select the range of numbers, sample size and the number of samples. When we increase the number
of sample to 1000s, the histogram starts looking like a normal graph which indicates that as the number
of sample tends to infinity, the mean of population is the mean of the range of numbers.  
  
When the user clicks on the submit button, the application  
1. Creates samples of sepcified size. It should be noted that the numbers are replaced during sampling and 
every number has equal chance of getting selected.  
2. Finds out the average of each sample.  
3. Creates the histogram of the averages.


## Code  
  
```{r echo = TRUE, eval=FALSE}
mean_pop <- reactive({
        range_min <- input$range[1]
        range_max <- input$range[2]
        averages <- range_min:range_max
        set.seed(666)
        for (i in 1:input$num_sample) {
            averages[i] <- mean(sample(range_min:range_max, 
                                       input$sample_size, 
                                       replace = TRUE))
        }
        averages <- data.frame(averages)
        return(averages)
    })
```
  
  
## Shiny Application
You may access the shiny application here:  
[**Shiny Application**](https://akshayamrit.shinyapps.io/CentralLimitTheorem/)  
  
To check out the code for the application, click here:  
[**GitHub Repository**](https://github.com/akshayamrit/CentralLimitTheorem.git) 