---
title: "Central Limit Theorem"
author: "Akshay Amrit"
date: "31st December, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objective  
This is a web page presentation created using R Markdown. This presentation is created as Final 
Peer Project of [**Developing Data Products**](https://www.coursera.org/learn/data-products/home/welcome) which is a part of [**Data Science Specialization**](https://www.coursera.org/specializations/jhu-data-science) by Johns Hopkins University on Coursera.  
  
The objective of this project is to prove Central Limit Theorem Experimentally.


## Process
User can select the range of numbers, sample size and the number of samples. When we increase the number
of sample to 1000s, the histogram starts looking like a normal graph which indicates that as the number
of sample tends to infinity, the mean of population is the mean of the range of numbers.  
  
When the user clicks on the submit button, the application  
1. Creates samples of sepcified size. It should be noted that the numbers are replaced during sampling.  
2. Finds out the average of each sample.  
3. Creates the histogram of the averages.


## Code  
  
```{r echo = TRUE, eval=FALSE}
mean_pop <- reactive({
        range_min <- input$range[1]
        range_max <- input$range[2]
        averages <- range_min:range_max
        set.seed(666)
        for (i in 1:input$num_sample) {
            averages[i] <- mean(sample(range_min:range_max, 
                                       input$sample_size, 
                                       replace = TRUE))
        }
        averages <- data.frame(averages)
        return(averages)
    })
```
  
  
## Shiny Application
You may access the shiny application here:  
[**Shiny Application**](https://akshayamrit.shinyapps.io/CentralLimitTheorem/)  
  
To check out the code for the application, click her:  
[**GitHub Repository**](https://github.com/akshayamrit/CentralLimitTheorem.git)  